---
title: "week12.Rmd"
author: "Kyrene"
date: "2026-04-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Script Settings and Resources
```{r}
set.seed(1234) # set seed for reproducibility
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
library(tidyverse)
library(tm)
library(qdap)
library(textstem)
library(tidytext)
# library(parallel) did not use
# library(doParallel) did not use
# library(tictoc) did not use
library(stm)
library(jsonlite)
library(caret)
# library(Matrix) did not use
```

# Data Import and Cleaning
```{r}
# I used this to download at least the last year of posts from r/IOPsychology.
# url <- "https://www.reddit.com/r/IOPsychology/new.json?limit=100"
# after <- NULL
# every_post <- list()
# cutoff_time <- Sys.time() - as.difftime(365, units = "days")
# done <- FALSE
# 
# repeat { # I use repeat() to create an infinite loop so it stops when I stop it with break. Keep requesitng pages and save each page until there are no more pages or reach to the point that it older than one year 
#   full_url <- paste0(url, if (!is.null(after)) paste0("&after=", after) else "")
#   res <- fromJSON(full_url)
#   posts <- res$data$children$data %>%
#     as_tibble()
#   
#   every_post[[length(every_post) + 1]] <- posts
#   after <- res$data$after
#   
#   oldest_post <- min(as.POSIXct(posts$created_utc, origin = "1970-01-01", tz = "UTC"))
#   if (is.null(after) || oldest_post <= cutoff_time) {
#     break
#   }
#   # this reduces the chance of rate limits
#   Sys.sleep(1)
# }
# 
# reddit_posts <- bind_rows(every_post)
# saveRDS(reddit_posts, "../data/reddit_posts.rds")
# 
# in the tibble, I only keep varibales that are required, convert unix time to normal time, keep posts from last year, and make sure it only has only upvotes and title
# week12_tbl <- reddit_posts %>%
#   select(title, upvotes = ups, created_utc) %>%
#   mutate(created_utc = as.POSIXct(created_utc, origin = "1970-01-01", tz = "UTC")) %>%
#   filter(created_utc > cutoff_time) %>%
#   select(upvotes, title)
# write_csv(week12_tbl, "../data/week12_tbl.csv")

# add code to import those files instead of downloading again

reddit_posts <- readRDS("../data/reddit_posts.rds")
week12_tbl <- read_csv("../data/week12_tbl.csv")
```

```{r, corpus}
# Create corpus from post titles, so later the original and preproccessed version can be compared.
io_corpus_original <- VCorpus(VectorSource(week12_tbl$title))

# Preprocessing using tm and qdap by building the transformer that removes references (IO). These terms appear frequently but do not distinguish the posts meaningfully. I think this is not too aggresive. regex might be more apprioproate here?
remove_io_terms <- content_transformer(function(x) {
  x %>%
    str_replace_all("\\bi\\s*/\\s*o\\s+psychology\\b", " ") %>%
    str_replace_all("\\bi\\s*/\\s*o\\s+psych\\b", " ") %>%
    str_replace_all("\\bi\\s*-\\s*o\\s+psychology\\b", " ") %>%
    str_replace_all("\\bi\\s*-\\s*o\\s+psych\\b", " ") %>%
    str_replace_all("\\bio\\s+psychology\\b", " ") %>%
    str_replace_all("\\bio\\s+psych\\b", " ") %>%
    str_replace_all("\\biopsychology\\b", " ") %>%
    str_replace_all("\\bindustrial\\s+and\\s+organizational\\s+psychology\\b", " ") %>%
    str_replace_all("\\bindustrial\\s+organizational\\s+psychology\\b", " ") %>%
    str_replace_all("\\bindustrial[- ]organizational\\s+psychology\\b", " ")
})

# I then created the preproccessed corpus. First, I expand abbreviations and contractions. Then, lowercase text so they are not case-sensitive. 
# Then, I use the cutome transformer to remove IO terms. Then, I remove numbers and punctuation becuase they are not that meaningful for post titles.
# Then, I lemmatize after punctuation & number removal, so words are simplified. I then remove stopwords after lemmatization. finally, I strip the extra whitespace at the end
io_corpus <- io_corpus_original %>%
  tm_map(content_transformer(replace_abbreviation)) %>%
  tm_map(content_transformer(replace_contraction)) %>%
  tm_map(content_transformer(str_to_lower)) %>%
  tm_map(remove_io_terms) %>%
  tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(content_transformer(lemmatize_strings)) %>%
  tm_map(removeWords, stopwords("en")) %>%
  tm_map(stripWhitespace)
```

```{r, compare them function}
# Function to compare the same randomly selected row from original corpus and the processed corpus.
# I repeatedly check whether preprocessing is doing its work cleanly
compare_them <- function(corpus_1, corpus_2) {
   i <- sample(1:length(corpus_1), 1)
  print(corpus_1[[i]]$content)
  print(corpus_2[[i]]$content)
}
compare_them(io_corpus_original, io_corpus)
```

```{r}
# Create a document id so each title stays linked to its row
io_text_tbl <- tibble(
  document = 1:length(io_corpus),
  text = sapply(io_corpus, as.character)
)

# Unigrams are used to capture individual high-frequency concepts.
io_unigram_tbl <- io_text_tbl %>%
  unnest_tokens(term, text, token = "words") %>%
  count(document, term)

# Bigrams are used ton capture multiword concepts
io_bigram_tbl <- io_text_tbl %>%
  unnest_tokens(term, text, token = "ngrams", n = 2) %>%
  count(document, term)

# Combine into one DTM called io_dtm and filter out any blank or missing terms that might result from preprocessing.
io_dtm <- bind_rows(io_unigram_tbl, io_bigram_tbl) %>%
  filter(!is.na(term), term != "") %>%
  group_by(document, term) %>%
  summarise(n = sum(n), .groups = "drop") %>%
  cast_dtm(document, term, n)

# Sparsity trimming
io_slim_dtm <- removeSparseTerms(io_dtm, .995)

# Check N/k ratio and adjust sparsity threshold to 0.995 to make it into the range 2:1 to 3:1
n_docs <- nrow(io_slim_dtm)
k_terms <- ncol(io_slim_dtm)
n_docs / k_terms
#  2.303867
```

# Visualization
```{r, Word cloud}
# I summrize total term frequencies across the full DTM and then produce a workdcloud. This can help me see what dominates discussion (only the words)
wordCounts <- slam::col_sums(io_dtm)
wordNames <- colnames(io_dtm)
wordcloud::wordcloud(wordNames, wordCounts, max.words = 50)
# interpret: the corpus is dominated by career/job-discussion content
```

# Analysis
```{r, topic analysis}
# I convert the dtm and then search for all possible topics. I then save the topic result plot into figs subfoolder, adjust the margins, draw the plot and close the graph so the file is actually written.
dfm2stm <- readCorpus(io_slim_dtm, type="slam")
kresult <- searchK(
  dfm2stm$documents,
  dfm2stm$vocab,
  K = seq(2, 20, by = 2)
)
png("../figs/kresult_plot.png", width = 1200, height = 900, res = 150)
par(mar = c(4, 4, 1, 1))
plot(kresult)
dev.off()

# I fit the final topic model with 6 topics.
# I choose 6 not 7 becuase 6 is better on semantic coherence and residuals 
# I then extract theta, the document-topic probability matrix and assigne it to topic_probs so each row is one document and each column is one topic
# For each document, I can identify topics with their highest probability and extract the probability of that most likely topics
topic_model <- stm(dfm2stm$documents, 
                   dfm2stm$vocab, 
                   6)
topic_probs <- topic_model$theta
assign_topic <- apply(topic_probs, 1, which.max)
assign_prob <- apply(topic_probs, 1, max)
keep_rows <- rowSums(as.matrix(io_slim_dtm)) > 0

# inspect the fitted topics using top words, representative titles, topic prevalence, and topic correlations
labelTopics(topic_model, n = 10)
findThoughts(topic_model, texts = week12_tbl$title[keep_rows], n = 3)
plot(topic_model, type = "summary", n = 6)
topic_correlation <- topicCorr(topic_model)
plot(topic_correlation)

# based on top words, I created labels and assign them into topic labels
topic_labels <- c(
  "1" = "Programs' Value & Degree values",
  "2" = "Research Opportunities & Helps",
  "3" = "People Analytics Careers",
  "4" = "Career Advice For Graduate",
  "5" = "Job Market",
  "6" = "Biweekly Discussion"
)

# Then, I created the tibble topics_tbl that has doc_id, original, topic, topic_label, probability, upvotes
topics_tbl <- tibble(
  doc_id = which(keep_rows),
  original = week12_tbl$title[keep_rows],
  topic = assign_topic,
  topic_label = topic_labels[as.character(assign_topic)],
  probability = assign_prob,
  upvotes = week12_tbl$upvotes[keep_rows]
)
```

# Explain in detail the specific reasoning process you used to determine the final number of topics 
I use searchK() to compare models across a range of topic numbers from K = 2 to K = 20 by 2
I start at 2 because K = 1 cannot meaningfully separate the posts into different themes
I then compare heldout likelihood, semantic coherence (semcoh), exclusivity (exclus), and residual-related measures
Based on these diagnostics, K = 6 gives the best overall balance between model fit and interpretability. It performsn better on semantic coherence and residuals and still keeps the topics reasonably distinct and easy to interpret
I therefore choose 6 because it separates the major themes well without making the topics too broad or too narrow

# Explain in detail the specific reasoning process for the topic labels you selected
I select the topic labels by looking at the top words in each topic, so I can understand what the main content in each topic is
I then check whether those top words match the actual post titles that are assigned to that topic
I use the top words and the titles together to figure out the main idea of each topic and then give each topic a short label to represent the idea
I also make the labels easy to understand, concise, and close to what the posts in each topic are about (2 to 5 words)

# Mahcine Learning
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# for machine leanring part, I first define the id and outcomes
ml_doc_id <- topics_tbl$doc_id
y <- topics_tbl$upvotes

# I create predictors matrxi use the token counts & create topic
x_tokens <- as.matrix(io_slim_dtm[ml_doc_id, ])
x_topics <- model.matrix( # turn those into a numeric matrix
  ~ factor(topic) + probability, # turn it into dummy variables and include the ptobability
  data = topics_tbl)[, -1] # remove the intercept
x_tokens_topics <- cbind(x_tokens, x_topics) # combine them so I can compare token alone against tokens-plus-topics

# split into training and holdout sets
train_index <- createDataPartition(topics_tbl$upvotes, p = .25, list = FALSE)[, 1]
train_tokens <- x_tokens[train_index, ] # token predictors for training
test_tokens <- x_tokens[-train_index, ] # token predictors for holdout
train_tokens_topics <- x_tokens_topics[train_index, ] # combine predictors for training
test_tokens_topics <- x_tokens_topics[-train_index, ] # combine predictors for holdout
train_y <- y[train_index] # upvotes for training
test_y <- y[-train_index] # upvotes for holdout

# fit the token-alone elastic net model
fit_tokens <- train(
  x = train_tokens,
  y = train_y,
  method = "glmnet",
  trControl = trainControl(method = "cv", number = 10, verboseIter = TRUE),
  preProcess = c("zv", "medianImpute", "center", "scale"),
  tuneGrid = expand.grid(alpha = c(0, 0.5, 1), lambda = seq(.001, .1, length.out = 10)),
  metric = "RMSE",
  na.action = na.pass
)


# fit the tokens-plus-topics model, this can isolate the effect of adding topic information
fit_tokens_topics <- train(
  x = train_tokens_topics,
  y = train_y,
  method = "glmnet",
  trControl = trainControl(method = "cv", number = 10, verboseIter = TRUE),
  preProcess = c("zv", "medianImpute", "center", "scale"),
  tuneGrid = expand.grid(alpha = c(0, 0.5, 1), lambda = seq(.001, .1, length.out = 10)),
  metric = "RMSE",
  na.action = na.pass
)

# I put all resampling results into one object, so it will be easier to compare cross-validated performance across models.
# Then, show summary statistics from the cross-validation results
resamples <- resamples(list(
  `Tokens only` = fit_tokens,
  `Tokens + topics` = fit_tokens_topics
))
summary(resamples)

# holdout predictions
pred_tokens <- predict(fit_tokens, newdata = test_tokens, na.action = na.pass)
pred_tokens_topics <- predict(fit_tokens_topics, newdata = test_tokens_topics, na.action = na.pass)

# final comparison table
table1_tbl <- tibble(
  model = c("Tokens only", "Tokens + topics"),
  cv_rsq = c(
    getTrainPerf(fit_tokens)$TrainRsquared,
    getTrainPerf(fit_tokens_topics)$TrainRsquared
  ),
  ho_rsq = c(
    postResample(pred_tokens, test_y)["Rsquared"],
    postResample(pred_tokens_topics, test_y)["Rsquared"]
  )
) %>%
  mutate(
    cv_rsq = str_remove(round(cv_rsq, 2), "^0"),
    ho_rsq = str_remove(round(ho_rsq, 2), "^0")
  )
write_csv(table1_tbl, "../out/table1.csv")
```

# Comment:
adding topics improves cv results but does not improve out-of-sample performance prediction. Both models perform poorly but the tokens plus topics model is slightly worse in the holdout set. CV improvement does not mean it can be generalized to new data.
